### 2013-07-29 sections/Acquisition.tex

\subsection{Data acquisition}
Describe the two-photon experiment, refer to Fig.\;1