### 2013-07-29 sections/Acquisition.tex
Added figure captions.

\subsection{Data acquisition}
Describe the two-photon experiment, refer to Fig.\;1

\textbf{Figure 1.}{
Experimental setup and data acquisition. 
\textbf{A}: Mouse in two-photon microscope, visual stimulation, 3D scanning of cells.
\textbf{B}: Calcium signals
\textbf{C}: Deconvolution and binning.
}\label{fig:01}