### 2013-08-08 sections/Acquisition.tex
updated the Theory section.

\subsection{Data acquisition}
Describe the two-photon experiment, refer to Fig.\;1.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/placeholder.pdf}
\caption{
%\textbf{Figure 1.}{
Experimental setup and data acquisition. 
\textbf{A}: Mouse in two-photon microscope, visual stimulation, 3D scanning of cells.
\textbf{B}: Calcium signals
\textbf{C}: Deconvolution and binning.
}\label{fig:01}
%}
\end{figure}