### 2013-08-09 sections/Simulation.tex
Added figure with simulation results.

To illustrate

\begin{figure}
\centering
\includegraphics[width=1.0\textwidth]{figures/Figure3.pdf}
\caption{
%\textbf{Figure 3.}{
Simulation results.
\textbf{A}: A diagonal matrix corresponds to a Gaussian graphical model with no dependencies. 
}\label{fig:03}
%}
\end{figure}