### 2013-08-10 sections/Simulation.tex

To illustrate

\begin{figure}
\centering
\includegraphics[width=1.0\textwidth]{figures/Figure3.pdf}
\caption{
%\textbf{Figure 3.}{
Simulation results.
\textbf{A--D}: True $60\times60$ covariance matrices used in models. \textbf{A} does not have a low-dimensional structure.  \textbf{B} is a factor model with 8 factors. \textbf{C}
}\label{fig:03}
%}
\end{figure}