### 2013-09-23 sections/Summary.tex
re-arranged sections according to PLOS guidelines

We estimated the covariance matrices of the calcium activity of populations of neurons in the primary visual cortex.  
In the effort to improve the estimate, we imposed several types of sparse structure of the covariance matrix and, using cross-validation, compared the performance for resulting estimators.