### 2013-09-25 sections/Figure1.tex
minor

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=4in]{figures/Figure1.pdf}
\end{center}
\caption{
{\bf Acquistion of neural population activity using two-photon fluorescence imaging of calcium signal.}  {\tt A.} Visual stimuli comprising brief (500 ms) presentatios of full-field drifting gratings separated by blank screens. {\tt B.} Two-photon fast 3D imaging of calcium signals in an awake mouse. {\tt C.} Deconvolved calcium signals. {\tt D.} The sample covariance matrix of neuronal calcium signals in 200 ms time bins. 
}
\label{Figure_label}
\end{figure}