### 2013-09-30 sections/Figure3.tex

\begin{figure}[htp]
\centering
\includegraphics[width=1.0\textwidth]{figures/Figure3.pdf}
\caption{
Selection amongst estimators recognizes covariance atructure.
}\label{fig:03}
\end{figure}