### 2013-09-30 sections/Figure5.tex

\begin{figure}[htp]
\centering
\includegraphics[width=0.5\textwidth]{figures/Figure2.pdf}
\caption{
Graphical models corresponding to the low-dimensional targets of the four regularization schemes used in the paper.
\textbf{A}: A diagonal matrix corresponds to a Gaussian graphical model with no dependencies.
\textbf{B}: In factor analysis, observed nodes are assumed to be influenced by several latent units (``factors") but are otherwise independent.
\textbf{C}: Graphical model with conditional dependencies between a specified subset of pairs of observed neurons, no latent units. 
\textbf{D}: Graphical model with conditional dependencies between a specified subset of pairs of observed neurons and dense interactions with a few latent units.
}\label{fig:02}
\end{figure}