### 2013-10-01 sections/Simulation.tex

\subsection*{In simulation, estimator selection `recognizes' the covariance structure}
\input{sections/Figure3.tex}