### 2013-10-01 sections/Figure5.tex

\begin{figure}[htp]
\centering
\includegraphics[width=1.0\textwidth]{figures/Figure5.pdf}
\caption{
Example of covariance structure revealed by the sparse+low-rank estimator and its relationship to the circuit arcitecture.
}
\label{fig:05}
\end{figure}