### 2013-10-01 sections/Findings.tex

\subsection*{In visual cortex, Estimator D dominates}
\input{sections/Figure4.tex}
Figure \ref{fig:04} showthe full matrix of all comparisons of the estimators.  The plots represents the histograms of the differences of the empirical loss between the pair of estimators for all 31 recorded sites.  When the differences are positive, the first estimate in the title outperforms the second estimate.

Of particular interest is the first row, which shows that all regularized estimates  outperformed the sample covariance estimate.  Even more informative is the last column, which shows that estimate D (sparse+lowrank) significantly outperformed all other estimates.