### 2013-10-02 sections/Simulation.tex

\subsection*{Simulation: estimators perform better when their target matches the low-dimensional structure of the true covariance matrix}
\input{sections/Figure3.tex}

A common misconception about regularization is that its effect depends on accurate prior knowlege about the structure of the data. However, substantial improvement can be attained by shrinking the unbiased estimate toward an arbitrary target as long as the target is less variable than the unbiased estimator.  This unintuitive phenomenon is known as \emph{Stein's paradox} \cite{Efron:1977}.  The more accurate description of regularization is the as the optimal tradeoff between estimation and approximation error, the so-called ``bias-variance tradeoff'' as described above.

Therefore, taken in isolation, the fact that a given regularized estimator improved the estimate should not be interpreted to suggest that its target estimate has the same low-dimensional structure as the data generating process.   However, when multiple estimators are compared against each other, the ones whose target estimate can closely approximate the true value with the smallest number of parameters will be able to reduce the estimation error with mininimum increases in approximation error when compared against other estimators whose targets cannot consistenly approximate the true covariance matrix.

Sample covariance matrices such as E,F,G,H are computed from samples of 1000 independent observations from multivariate normal distributions with true covariance matrices in A,B,C,D, respectively.  However, the true covariance matrices have various low-dimensional structures.

True matrix A has no low-dimensional structure.  It is computed as the sample covariance matrix of a finite sample (n=1000) of a multivariate Gaussian with independent identically distributed nodes.

True matrix B is computed by factor analysis of true matrix A with six factors.

True matrix C is computed by reducing to zero a large fraction the coefficients of the inverse of true matrix A.

True matrix D is computed by representing the inverse of true matrix A as the sum of a sparse component and a low-low rank component.

When we estimate the true covariance matrix from the samples of various sizes, we find that for true covariance matrices that have a low-dimensional structure, the covariance estimator whose target has the same type of low-dimensional structure attains lower loss. Plots I,J,K,L show the excess loss, which equals zero when the estimate equals truth (requires knowledge of truth). Plots M,N,O,P show empirical loss computed by cross-validation.  Empirical loss does not require knowledge of truth but is also shifted so that its absolute value is not particularity meaningful but it can be used for comparison. The error bars in these plots indicate the standard deviation, not standard error.

The loss function here is the multivariate log likelihood just as used in the previous post on real data.

The fact that estimate D outperformed the other estimates can be interpreted that the low-dimensional structure of the covariance matrices obtained from dense microcircuits in the visual cortex is best represented by the sparse+low-rank inverse model.

We can be then justified to examine the optimal fitted model rather than the sample covariance model in the analysis of neural correlations.  The sparse component represents the graph of interactions in the model, which we can now relate to the circuit architecture.  The low-rank component can be interpreted as common input into the circuit whose properties can be studied separately.

In this way, the process of selecting the best of several regularized covariance estimators serves as the method for recognizing the low-dimensional structure of covariance matrices in a specific domain.

Note that estimator D performs well on both purely sparse and purely low-rank models. But it does not do better.  Only when the truth is sparse+lowrank does it outperform the other estimates.