### 2013-10-04 sections/Figure5.tex
added contents to Fig. 1 and Fig. 5

\begin{figure}[htp]
\centering
\includegraphics[width=1.0\textwidth]{figures/Figure5.pdf}
\caption{
Example of covariance structure revealed by the sparse+low-rank estimator and its relationship to the circuit arcitecture.
{\bf A} The correlation matrix corresponding to the sample covariance matrix in Fig.~\ref{fig:01}D. {\bf B} The inverse of the sample covariance matrix (sample concentration matrix). {\bf D} shows the result of covariance matrix estimation by estimator $\mathcal D$ and {\bf E} show its inverse. The sparse ({\bf C}) and low-rank ({\bf F}) components of the inverse.  {\bf G} depicts the spatial and functional organization of the connections represented by the sparse component of the estimated concentration matrix. 
}
\label{fig:05}
\end{figure}