### 2013-10-23 sections/Figure4.tex
redid figure 4 to show only estimator D.

\begin{figure}[htp]
\centering
\includegraphics[width=0.5\textwidth]{figures/Figure4.pdf}
\caption{Estimator D dominates over the other estimators in estimation of noise covariances in dense recordings from primary visual cortex.
}\label{fig:04}
\end{figure}