### 2013-10-24 sections/Figure4.tex

\begin{figure}[htp]
\centering
\includegraphics[width=0.5\textwidth]{figures/Figure4.pdf}
\caption{{\bf Estimator $\mathcal D$ (sparse+full-rank inverse covariance) dominates in dense recordings from primary visual cortex.}
{\bf A:}  Histogram of improvements in validation loss by covariance estimator $\mathcal D$ relative to the samle covariance matrix from the 31 imaged sites used in the study.  The vertical red line denotes the median improvment. 
{\bf B:} Improvements relative to estimator $\mathcal A$ (shrinkage toward the diagonal).
{\bf C:} Improvements relative to estimator $\mathcal B$ (shrinkage toward the factor model).
{\bf D:} Improvements relative to estimator $\mathcal C$ (sparse inverse covariance).
}\label{fig:04}
\end{figure}